KIN: A method to infer relatedness from low-coverage ancient DNA

### 2022-04-29 paper_overleaf/main.tex

For detecting ROH, there is only one method available that works with low coverage ancient DNA \cite{ringbauer_parental_2021}. This software can infer ROH at coverages $>0.3x$, but requires a reference panel. In the absence of a method to detect ROH at low coverages without a reference panel to detect long ROH, and we tested it on simulations to show that it can be used reliably at coverage $>=0.1x$. 
In our simulations, we find that adding long ROH ($~17\%$) to simulations slightly improves the power of both KIN and READ, particularly at lower coverages. This is likely because ROH actually reduces the variance in differences, and so it becomes easier to estimate the location of IBD tracts. For READ, the presence of ROH causes a bias towards inferring closer relatedness cases, but the model we use in KIN reduces this bias (see Fig. \ref{fig3:Comparison_READ_KIN}).

@@

\section{Acknowledgements}
We thank Svante Pääbo, Janet Kelso, Harald Ringbauer, Fabrizio Mafessoni, Laurits Skov, Leonardo Iasi, Alba Bossom Mesa, Arev Sümer and Luca Traverso for helpful comments and discussions. This work was supported by the Max Planck Society and the European Research Council (Grant No.: 694707) to Svante Pääbo.

>>>>

For detecting ROH, there is only one method available that works with low coverage ancient DNA \cite{ringbauer_parental_2021}. This software can infer ROH at coverages $>0.3x$, but requires a reference panel. In the absence of a method to detect ROH at low coverages without a reference panel, and we tested it on simulations to show that it can be used reliably at coverage $\geq 0.1x$ to detect ROH regions at lengths on the order of 10cM. 
In our simulations, we find that adding long ROH ($~17\%$) to simulations slightly improves the power of both KIN and READ, particularly at lower coverages. This is likely because ROH actually reduces the variance in differences, and so it becomes easier to estimate the location of IBD tracts. For READ, the presence of ROH causes a bias towards inferring closer relatedness cases, but the model we use in KIN reduces this bias (see Fig. \ref{fig3:Comparison_READ_KIN}).

@@

\section{Acknowledgements}
We thank Svante Pääbo, Janet Kelso, Harald Ringbauer, Laurits Skov, Leonardo Iasi, Alba Bossom Mesa, Arev Sümer, Guido Alberto Gnecchi Ruscone, Ke Wang and Luca Traverso for helpful comments and discussions. This work was supported by the Max Planck Society and the European Research Council (Grant No.: 694707) to Svante Pääbo.
